MLH1 Results Outline

### 2020-02-19 doc/Results.rmd
knited a new version

-We focus on observations from two-crossover bivalents for more comparable observations.

-comparisons of all IFDs of mulit crossover bivalents is also included

Interference is a major determinant of the positioning of chromosomes

@@

```{r show.IFD_99, echo=FALSE, warning=FALSE}

@@

```{r IFD.old.plots, echo=FALSE, eval=FALSE}
#long SC Long_biv_mouse.avs_4MM
#real.long.bivData
#Only use 2CO bivalents
real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]

@@

Still working on the best way to display the general IFD patterns.

@@

## Strain Comparisons

@@

# Q2 Analysis Predictions, Male Polymorphism and (High vs Low Rec strains)

<!-- keep blank for now 
-basic stats for the new dataset

@@

whoo.mama <-  t.test( (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)  %>%  filter(Rec.group == "0"))$PER_Foci_1,

            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   %>% filter(Rec.group == "1"))$PER_Foci_1   )

>>>>

-We focus on observations from two-crossover bivalents for more comparable observations.
-comparisons of all IFDs of mulit crossover bivalents is also included
Interference is a major determinant of the positioning of chromosomes

@@

<!--  set up -->

<!--  plots-->

```{r Q1.IFD_show.IFD_99, echo=FALSE, warning=FALSE}

@@

```{r Q1.IFD_IFD.old.plots, echo=FALSE, eval=FALSE}
#long SC Long_biv_mouse.avs_4MM
#real.long.bivData
#Only use 2CO bivalents
real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]

@@

<!--t.tests -->

```{r Q1.IFD_Ttests, echo=FALSE}

#below are calq just from 2COs, have no fear
#mean_IFD.2CO_ABS
#mean_IFD.2CO_PER

#FULL all strains
Q1.IFD.raw_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_ABS,

                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_ABS)
#p-value = 0.07

Q1.IFD.PER_FULL_t.test <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(strain !=  "CZECH" ) %>% filter(strain !=  "MOLF" )    )$mean_IFD.2CO_PER,

                                ( mouse.avs_4MM %>% filter(sex == "female") )$mean_IFD.2CO_PER)
#p-value = 1e-12

The raw IFD measures for 2CO bivalents are slightly significant (t.test; p = `r Q1.IFD.raw_FULL_t.test$p.value`). While the normalized IFD measures are highly significant (t.test; p = `r Q1.IFD.PER_FULL_t.test$p.value`)

(interpertation statement)  Males have stronger interference

<!-- lmer/ glms-->

@@

@@

@@

whoo.mama <-  t.test( (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1)  %>%  filter(Rec.group == "0"))$PER_Foci_1, 
            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   %>% filter(Rec.group == "1"))$PER_Foci_1   )
#p-value = 0.2

whoo.mama.rec.exc <-  t.test( (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)  %>%  filter(Rec.group.exclu == "0"))$PER_Foci_1, 
            (Curated_BivData  %>% filter(sex == "male") %>% filter(hand.foci.count == 1)   %>% filter(Rec.group.exclu == "1"))$PER_Foci_1   )

#p-value = 0.4

The t.test at the bivalent level is also NS (test; p = )

```{r Q2.1COpos_within.subsp_ttest, echo=FALSE}

#compare within subsp

#Musc
Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
            (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1"))$norm.first.foci.pos   )

# p-value = 0.4
# exclusive rec group not sig
```

**Most all of the t.tests I've done indecate that there are no sig diff between the High and the low rec groups**

The normalized 1CO position is not a good metric to distinguish the high and low rec groups.

GLMs

glms are constructed to test if there has been evolution between strains ... ect.
I will use the full data set the reduced datasets have such low power

```{r Q2.1COpos_mus.av.full_glm, echo=FALSE}

#(mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos

Q2.glm_mus.av_1COpos_M1 <- glm(norm.first.foci.pos ~ subsp * strain, data= mouse.avs_4MM %>% filter(sex == "male"))

sum.Q2.glm_mus.av_1COpos_M1  <- summary(Q2.glm_mus.av_1COpos_M1)$coefficients

#M2
Q2.glm_mus.av_1COpos_M2 <- glm(norm.first.foci.pos ~ strain, data= mouse.avs_4MM %>% filter(sex == "male"))

sum.Q2.glm_mus.av_1COpos_M2 <- summary(Q2.glm_mus.av_1COpos_M2)$coefficients

for M1
Mol subsp is significant, strain G (LEW slight sig), strain MSM -- strain MOLF is NA

for M2,

G, LEW, PWD, MSM, MOLF, SKIVE, KAZ have significant effects

put together, these glm's indeceate that MOLF is the most different in it's 1CO landscape (compared to other males). Is this a sampling issue

For MOLF 326 1CO bivalents,  79, 126, 62, 59 1CO observations across the 4 MOLF mice

there are only 77 MSM 1COs across 4 mice in MSM (MSM is less central than MOLF)
(motivation for MOLF female observations and more male MSM biv measures)
